BART: A Comparison with Non-bayesian Tree Methods

### 2024-12-07 paper/paper.tex
Adding section of paper

\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{color}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage{booktabs}
\usepackage{caption}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subcaption}
\usepackage{bookmark}
\usepackage[
    backend=biber,
    style=authoryear,
  ]{biblatex}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{listing}
\usepackage{color}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage{booktabs}
\usepackage{caption}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subcaption}
\usepackage{bookmark}
\usepackage[
    backend=biber,
    style=authoryear,
  ]{biblatex}

@@

\section{Data Application: Predicting the age of an Abalone}

Abalone are marine snails found in various regions around the world. Determining their age is a labor-intensive process, as it requires cutting through the shell, staining it, and then counting the growth rings under a microscope. However, predicting the age of an abalone can also be done by examining other physical characteristics of the individual, which are much easier to assess. This is precisely the motivation behind our application.

We use data Abalone characteristics data coming from the UC Irvine Machine Learning Repository \parencite{warwicknashAbalone1994a}, that is described in table \ref{table1}. No data manipulation is necessary, aside from creating a new variable, \( \text{Age} = \text{Rings} + 1.5 \), and removing the Rings column.

 We run BART from \cite{mccullochBARTBayesianAdditive2024}; Boosting from \cite{ridgewayGbmGeneralizedBoosted2024}; Random Forest and Bagging from \cite{breimanRandomForestBreimanCutlers2024}. We use half of the sample to fit $\hat{y}$, that is the Age variable, and half the sample to give us out of bag Mean Squared Errors (MSE) for predicting the same variable. Models are run for various values of TT, and for each value, both the mean squared error (MSE) of out-of-sample predictions (shown in plot \ref{plot_mse}) and the computation time (depicted in plot \ref{plot_time}) are recorded.

\begin{table}[]
  \centering
  \begin{tabular}{llllll}

  \toprule
  Variable Name & Role & Type & Description & Units & Missing Values\\  
  \midrule
   Sex                  & Feature       & Categorical   & M, F, and I (infant) & -             & No                      \\   Length               & Feature       & Continuous    & Longest shell measurement & mm         & No                      \\   Diameter             & Feature       & Continuous    & Perpendicular to length & mm           & No                      \\   Height               & Feature       & Continuous    & With meat in shell      & mm           & No                      \\   Whole\_weight         & Feature       & Continuous    & Whole abalone           & grams        & No                      \\   Shucked\_weight       & Feature       & Continuous    & Weight of meat          & grams        & No                      \\   Viscera\_weight       & Feature       & Continuous    & Gut weight (after bleeding) & grams    & No                      \\   Shell\_weight         & Feature       & Continuous    & After being dried       & grams        & No                      \\   Rings                & Target        & Integer       & +1.5 gives the age in years & -          & No                      \\
   \bottomrule
  \end{tabular}
  \caption{Table of Variables}
  \label{table1}
  \end{table}

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{../outputs/mse_plot.pdf}
  \caption{Mean Squared Error for different values of T}
  \label{plot_mse}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{../outputs/time_plot.pdf}
  \caption{Time of computations for different values of T}
  \label{plot_time}
\end{figure}
